---
title: "datamining some cbf data"
author: "Einar Hjörleifsson"
date: "25 October 2017"
output: 
  html_document: 
    fig_caption: yes
    code_folding: hide
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
library(stringr)
library(viridis)
library(broom)
library(modelr)
```

```{r}
fil <- "CBF_shared/Session plans/Upali-1-1a-Exercise-worksheet.xlsx"
d1 <- read_excel(fil,
                 range = "A2:G47")
colnames(d1) <- c("id", "sdd", "tp", "chla", "y", "sd", "redundant")
d1 <-
  d1 %>% 
  select(-redundant) %>% 
  separate(sdd, c("sdd", "ssd_sd"), sep = " ± ", convert = TRUE) %>% 
  separate(tp,  c("tp", "tp_sd"), sep = " ± ", convert = TRUE) %>% 
  separate(chla, c("chla", "chla_sd"), sep = " ± ", convert = TRUE) %>% 
  mutate(country = "Sri Lanka",
         id = as.integer(str_sub(id, 2)))
d2 <- 
  read_excel(fil,
             range = "A50:F73")
colnames(d2) <- c("id", "area", "shoreline", "rla", "dl", "y")
d2 <-
  d2 %>% 
  mutate(id = as.integer(str_sub(id, 2)))
d1 <-
  d1 %>% 
  left_join(d2 %>% select(-y))
d2$country <- "Sri Lanka"
d3 <- 
  read_excel(fil,
             range = "A76:H95")
colnames(d3) <- c("id", "name", "area", "ssd", "y", "sd", "wgt", "dummy")
d3 <- 
  d3 %>% 
  select(-dummy) %>% 
  mutate(country = "Bangladesh")
d4 <- 
  read_excel(fil,
             range = "A98:F138")
colnames(d4) <- c("id", "name", "area", "sd", "y", "efficiency")
d4$country <- "India"
d5a <- 
  read_csv("inst/csv/vietnam_stocking_density.csv") %>% 
  gather(year, sd, c(-name, -area), convert = TRUE) %>% 
  mutate(sd = sd/0.025)
d5b <- 
  read_csv("inst/csv/vietnam_yield.csv") %>% 
  gather(variable, value, -name) %>% 
  separate(variable, c("variable", "year"), convert = TRUE) %>% 
  spread(variable, value) %>% 
  rename(y = yield)
d5 <-
  d5a %>% 
  full_join(d5b) %>% 
  mutate(country = "Vietnam") %>% 
  filter(!is.na(sd), !is.na(y))

d <-
  bind_rows(d1, d3) %>% 
  bind_rows(d4) %>% 
  bind_rows(d5)
```

```{r}
# Estimating the optimum stocking density

# The model
country_model <- function(df) {
  nls(y ~ a * sd + b * sd^2,
      data = df, start=list(a = 0, b = 2))
}
# Estimate
d2 <-
  d %>% 
  group_by(country) %>% 
  nest() %>% 
  mutate(model = map(data, country_model)) %>% 
  mutate(pre = map2(data, model, add_predictions),
         res = map2(data, model, add_residuals))
# A trick to get a continuous line on plot
pre <- 
  expand.grid(sd = c(seq(0, 12000, by = 50), seq(15000, 44000, by = 1000)),
              country = unique(d$country),
              stringsAsFactors = FALSE) %>%
  as_tibble() %>% 
  group_by(country) %>% 
  nest() %>% 
  left_join(d2 %>% select(country, model)) %>% 
  mutate(pred = map2(data, model, add_predictions)) %>% 
  unnest(pred) %>% 
  filter(pred >= 0)
# Communication part
ggplot() +
  geom_point(data = d, aes(sd, y, colour = country)) +
  geom_line(data = pre, aes(sd, pred, colour = country)) +
  labs(x = "Stocking density [seedlings/hectare]",
       y = "Yield [kg/hectare]",
       colour = "Country",
       title = "Yield as a function of stocking density by country") +
  theme(legend.position = c(0.8, 0.7)) +
  scale_color_brewer(palette = "Set1")
```


```{r}
d %>% 
  ggplot(aes(area, y, colour = country)) +
  geom_point() +
  theme(legend.position = c(0.8, 0.7)) +
  scale_color_brewer(palette = "Set1")
```

```{r}
d2 %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  unnest(glance, .drop = TRUE) %>% 
  knitr::kable()
```

### A. Evaluation of trophic state of reservoirs.

Carlon’s Trophic State Index (TSI) can be estimated using total phosphorous (TP), chlorophyll-a content (Chl-a) and Secchi disc depth (SDD).

1) Estimate TSI (TP), TSI (Chl-a) and TSI (SDD) for the individual reservoirs using the following equations.

* TSI (TP) = 14.42 ln Total P (mg/m3) + 4.15
* TSI (Chl-a) = 9.81 ln Chlorophyll a (mg/m3) + 30.6
* TSI (SDD) = 60 – 14.41 ln SDD (metres)

2) Estimate average TSI for individual reservoirs as follows:

* Average TSI = [TSI(TP) + TSI(Chl-a) + TSI(SDD)]/3

```{r}
d <-
  d %>% 
  mutate(tsi_tp = 14.42 * log(tp) + 4.15,
         tsi_chla = 9.81 * log(chla) + 30.6,
         tsi_sdd = 60 - 14.41 * log(sdd),
         tsi = (tsi_tp + tsi_chla + tsi_sdd) / 3)
```

3) Using the following trophic classification of lakes and reservoirs based on Carlson’s TSI, classify each reservoir.

```{r, eval = FALSE}
tcl <- read_csv("classificatoin")
tcl2 <- read_excel("tci.xlsx")
knitr::kable(tcl)
x <-  
  mutate(class = case_when(d1$tsi < 30 ~ "<30"))
```

4) Determine the relationship between TSI and CBF yield in **23** reservoirs, using linear regression analysis.

Comment on the Trophic State and CBF potential in non-perennial reservoirs.

```{r}
d %>% 
  select(tsi_sdd, tsi_tp, tsi_chla,tsi, y) %>% 
  gather(variable, value, -y) %>% 
  ggplot(aes(value, y)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ variable, scale = "free") +
  labs(x = NULL, y = "Yield")
```

```{r}
d %>% 
  ggplot(aes(area, y)) +
  geom_point(aes(colour = country)) +
  #geom_smooth(aes(colour = country), span = 1 ) +
  labs(x = "Area", y = "Yield") +
  coord_cartesian(ylim = c(0,1800)) +
  theme(legend.position = c(0.8, 0.7))
```





```{r}
d <- 
  d %>% 
  mutate(cpus = y/sd)
d %>% 
  ggplot(aes(sd, cpus)) +
  geom_point(aes(colour = country), alpha = 0.4) +
  expand_limits(x = 0, y = 0)

fit <- lm(cpus ~ sd + area, data = d %>% filter(country == "Sri Lanka"))
summary(fit)
fit <- lm(cpus ~ sd + area, data = d %>% filter(country == "India"))
summary(fit)
fit <- lm(cpus ~ sd + area, data = d %>% filter(country == "Bangladesh"))
summary(fit)
fit <- lm(cpus ~ sd + area, data = d %>% filter(country == "Vietnam"))
summary(fit)


fit.full <- lm(cpus ~ sd + area + as.factor(country), data = d)
summary(fit.full)
drop1(fit.full)
fit.reduced <- step(fit.full)
fit.reduced

d1 <-
  d %>% 
  filter(country == "Sri Lanka")
fit <- lm(cpus ~ sd + area + sdd + tp + chla, data = d1)
summary(fit)
drop1(fit)
fit.reduced <- step(fit)
```



# Extra

```{r}
d %>% 
  ggplot(aes(sd, wgt, colour = country)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
d %>% 
  filter(country == "Bangladesh") %>% 
  ggplot(aes(sd, y)) +
  #geom_point(aes(size = wgt)) +
  geom_text(aes(label = wgt), angle = 45) +
  scale_size() +
  scale_colour_viridis(option = "B", direction = -1) +
  expand_limits(x = 0, y = 0)
```

```{r}
a <- 2.5e-5
b <- 3
l2w <- function(l, a = 2.5e-5, b = 3) {
  a * l^b
}
w2l <- function(w, a = 2.5e-5, b = 3) {
  exp((log(w) - log(a)) / b)
}

d3 <-
  d3 %>% 
  mutate(length = w2l(wgt),
         n = y/wgt,
         m = n/sd)
d3 %>% 
  ggplot(aes(length)) +
  geom_histogram()
d3 %>% 
  ggplot(aes(sd, m)) +
  geom_point() +
  geom_smooth(method = "lm")
```


```{r}
d %>% 
  ggplot(aes(sd, y, colour = country)) +
  geom_point()
```


