---
title: "cbf - exercises"
author: "Einar Hjörleifsson"
date: "25 October 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
```


```{r}
d1 <- read_excel("CBF_shared/Session plans/Upali-1-1a-Exercise-worksheet.xlsx",
                 range = "A2:G47")
colnames(d1) <- c("id", "sdd", "tp", "chla", "y", "dens", "redundant")
d1 <-
  d1 %>% 
  select(-redundant) %>% 
  separate(sdd, c("sdd", "ssd_sd"), sep = " ± ", convert = TRUE) %>% 
  separate(tp,  c("tp", "tp_sd"), sep = " ± ", convert = TRUE) %>% 
  separate(chla, c("chla", "chla_sd"), sep = " ± ", convert = TRUE)
```

## Evaluation of trophic state of reservoirs.

Carlon’s Trophic State Index (TSI) can be estimated using total phosphorous (TP), chlorophyll-a content (Chl-a) and Secchi disc depth (SDD).

### Estimate TSI (TP), TSI (Chl-a) and TSI (SDD) for the individual reservoirs using the following equations.

* TSI (TP) = 14.42 ln Total P (mg/m3) + 4.15
* TSI (Chl-a) = 9.81 ln Chlorophyll a (mg/m3) + 30.6
* TSI (SDD) = 60 – 14.41 ln SDD (metres)

```{r}
d1 <-
  d1 %>% 
  mutate(tsi_tp = 14.42 * log(tp) + 4.15,
         tsi_chla = 9.81 * log(chla) + 30.6,
         tsi_sdd = 60 - 14.41 * log(sdd))
```

### Estimate average TSI for individual reservoirs as follows:

* Average TSI = [TSI(TP) + TSI(Chl-a) + TSI(SDD)]/3

```{r}
d1 <- 
  d1 %>% 
  mutate(tsi = (tsi_tp + tsi_chla + tsi_sdd) / 3)
```

### Using the following trophic classification of lakes and reservoirs based on Carlson’s TSI, classify each reservoir.

```{r}
trophic_classfication <- read_csv("inst/csv/trophic_classification.csv")
knitr::kable(trophic_classfication)
```

### Determine the relationship between TSI and CBF yield in **23** reservoirs, using linear regression analysis.

```{r}
d1 %>% 
  ggplot(aes(tsi, y)) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method = "lm") +
  coord_cartesian(ylim =  c(0, 2000))
```


Comment on the Trophic State and CBF potential in non-perennial reservoirs.

# speater
```{r}
d2 <- 
  read_excel("CBF_shared/Session plans/Upali-1-1a-Exercise-worksheet.xlsx",
                 range = "A50:F73")
colnames(d2) <- c("id", "area", "shoreline", "rla", "dl", "y")
```

